### 2014-10-18 slides/slides.tex
Add slides folder

%
% $Id: $
%
%
% Compilar a .pdf con LaTeX (pdflatex)
% Es necesario instalar Beamer (paquete latex-beamer en Debian)
%

%
% Gráficos:
% Los gráficos pueden suministrarse en PNG, JPG, TIF, PDF, MPS
% Los EPS deben convertirse a PDF (usar epstopdf)
%

\documentclass{beamer}
\usetheme{Warsaw}
%\usebackgroundtemplate{\includegraphics[width=\paperwidth]{format/libresoft-bg.png}}
%\usepackage[spanish]{babel}
\usepackage[latin1]{inputenc}
\usepackage{graphics}
\usepackage{amssymb} % Simbolos matematicos
\usepackage{url}

%\definecolor{libresoftgreen}{RGB}{162,190,43}
%\definecolor{libresoftblue}{RGB}{0,98,143}

%\setbeamercolor{titlelike}{bg=libresoftgreen}

%% Metadatos del PDF.
\hypersetup{
  pdftitle={Automatic Detection of Bad Programming Habits in Scratch: A Preliminary Study},
  pdfauthor={Jesús Moreno León, Gregorio Robles},
  pdfcreator={GSyC/LibreSoft \\ Universidad Rey Juan Carlos},
  pdfproducer=PDFLaTeX,
  pdfsubject={Learning to code with Scratch. Automatic assesment.},
}
%%

\begin{document}

\title{Automatic Detection of Bad Programming Habits in Scratch}
\subtitle{A Preliminary Study}
\institute{jesus.moreno@programamos.es, grex@gsyc.urjc.es \\
GSyC/Libresoft, Universidad Rey Juan Carlos}
\author{Jesús Moreno León, Gregorio Robles}
\date{FIE 2014, Madrid, October 23 2014}

\frame{
\maketitle
\begin{center}
\includegraphics[width=2cm]{format/libresoft-logo}
\hspace{0.5cm}
\includegraphics[width=5cm]{format/gsyc-urjc}
\vspace{0.5cm}
\includegraphics[width=3cm]{format/emadrid.png}
\end{center}
}

% Si el titulo o el autor se quieren acortar para los pies de página
% se pueden redefinir aquí:
%\title{Titulo corto}
%\author{Autores abreviado}

%% LICENCIA DE REDISTRIBUCION DE LAS TRANSPAS
\frame{
~
\vspace{3cm}

\begin{flushright}
\includegraphics[width=2.2cm]{figs/by-sa}

{\tiny
(cc) 2014 Gregorio Robles and Jesús Moreno León\\
  Some rights reserved. This work licensed under Creative Commons \\
  Attribution-ShareAlike License. To view a copy of full license, see \\
  http://creativecommons.org/licenses/by-sa/3.0/ or write to \\
  Creative Commons, 559 Nathan Abbott Way, Stanford, \\
  California 94305, USA. \\
\ \\
Some of the figures have been taken from the Internet \\
Source, and author and licence if known, is specified. \\
For those images, \emph{fair use} applies.
}
\end{flushright}
}
%%

\section{About}

%--------------------------------------------------------
\usebackgroundtemplate{\includegraphics[width=19cm]{figs/AngryBirds.png}}

\begin{frame}
\frametitle{Goal of our paper}

\begin{center}
\Huge Are bad programming habits a common practice in the Scratch community?
\end{center}

\end{frame}

%-----------------------    ---------------------------------
\usebackgroundtemplate{\includegraphics[width=14cm]{figs/audience.png}}
% http://cdn.netrafic.com/wp-content/uploads/2011/02/audience.gif

\begin{frame}
\frametitle{Audience}

Who should/could be interested in this talk?

\begin{itemize}
  \item Educators teaching to code
  \item Students learning to code
  \item Programming learning tools developers
\end{itemize}

\end{frame}

\usebackgroundtemplate{}
%--------------------------------------------------------
\begin{frame}
\frametitle{Scratch}

  \begin{columns}[T]
    \begin{column}{1\textwidth}
     \begin{block}{Learning to code with Scratch}
\begin{itemize}
  \item Success to teach both basic and advanced programming concepts
  \item Bad programming habits detected
  \item Lack of automatic tools to check for \textit{correctness}
  \item Hairball: Lint-inspired Static Analysis of Scratch Projects
\end{itemize}
    \end{block}
    \end{column}

  \end{columns}

\end{frame}

\usebackgroundtemplate{}

%--------------------------------------------------------
\begin{frame}
\frametitle{Bad programming habits with Scratch: objects naming}

\begin{figure}[t!]
\begin{center}
\includegraphics[width=7cm, height=6cm]{figs/SpriteNaming.png}
\end{center}
\label{fig:naming}
\end{figure}

\begin{center}
Sprites naming
\end{center}
\end{frame}

%--------------------------------------------------------
\begin{frame}
\frametitle{Bad programming habits with Scratch: repetition code}

\begin{figure}[t!]
\begin{center}
\includegraphics[width=6cm,height=2.7cm]{figs/CodeRepetition1.png}
\end{center}
\label{fig:repetition1}
\end{figure}

\begin{center}
Repetition code
\end{center}

\begin{figure}[t!]
\begin{center}
\includegraphics[width=6cm,height=2.7cm]{figs/CodeRepetition2.png}
\end{center}
\label{fig:repetition2}
\end{figure}

\begin{center}
Creating blocks to avoid repetition code
\end{center}

\end{frame}

%--------------------------------------------------------
%--------------------------------------------------------
\usebackgroundtemplate{\includegraphics[width=13cm,height=9.2cm]{figs/plugins.png}}
% background: http://25.media.tumblr.com/b83aa72682992ab34b8ce7e61c0cb7f9/tumblr_menxc7qcq61ryin08o1_r1_1280.jpg
\begin{frame}
\frametitle{Hairball plug-ins development}

We developed two plug-ins for Hairball to automatically detect bad programming habits:
\begin{itemize}
  \item convention.SpriteNaming
  \item duplicate.DuplicateScripts
\end{itemize}

\end{frame}

\usebackgroundtemplate{}

%--------------------------------------------------------
%\usebackgroundtemplate{\includegraphics[width=13cm]{figs/iceberg.jpg}}
% background: http://www.wim-network.org/wp-content/uploads/2012/04/iceberg.jpg

\begin{frame}
\frametitle{Scratch projects repository analysis}

\begin{table}
\begin{center}
  \begin{tabular}{ | l | r | r | r |}
   \hline
              & Default names & Duplicated scripts & Defined blocks \\ \hline\hline
    Projects & 79 & 62 & 17 \\ \hline
    Mean & 5.94 & 7.23 & 1.11 \\ \hline
    Median & 3 & 2 & 0 \\ \hline
    Maximum & 67 & 71 & 25 \\
    \hline    
  \end{tabular}
\end{center}
\caption{Analysis of 100 ramdonly downloaded projects}
\label{table:results}
\end{table}
\end{frame}

%--------------------------------------------------------
%\usebackgroundtemplate{\includegraphics[width=13cm]{figs/take-away.jpg}}
% background: http://2.bp.blogspot.com/-78Eh4TBpdtU/UPw7ULV73PI/AAAAAAAAHAE/6DQfvPNCo-Y/s1600/8723052-stylized-red-stamp-showing-the-term-take-away-all-on-white-background.jpg
\usebackgroundtemplate{\includegraphics[width=13cm]{figs/future.png}}

\begin{frame}
\frametitle{Future Work}

\begin{itemize}
  \item Extend the scope of the study developing new plug-ins
  \item Analyze dataset with 5 years data from Scratch website
  \item Dr.Scratch: http://drscratch.programamos.es
\end{itemize}
\vspace{\baselineskip}
\vspace{\baselineskip}
\hfill{\Tiny Background picture: Simon Cunningham }

\end{frame}

%--------------------------------------------------------
%\begin{frame}
%\frametitle{GSyC/LibreSoft}

%\begin{figure}[t!]
%\begin{center}
%\includegraphics[width=11cm]{figs/libresoft.jpg}
%\end{center}
%\label{fig:libresoft}
%\end{figure}

%\begin{center}
%The GSyC/LibreSoft research team at URJC (Madrid).
%\end{center}

%\end{frame}

%--------------------------------------------------------
%\begin{frame}
%\frametitle{Our work at GSyC/LibreSoft}

%\begin{figure}[t!]
%\begin{center}
%\includegraphics[width=2.5cm]{figs/research}
% http://www.memphis.edu/crow/images/research_2.jpg
%\hspace{0.1cm}
%\includegraphics[width=2.5cm]{figs/teaching}
% http://2.bp.blogspot.com/_uxgwfriLwSo/TOWDr8IjaLI/AAAAAAAABLM/d7H-G5jIq-c/s1600/teaching.gif
%\hspace{0.1cm}
%\includegraphics[width=2.5cm]{figs/development}
% http://www.vidadigitalradio.com/wp-content/uploads/2009/04/hackers_cartoons.jpg
%\hspace{0.1cm}
%\includegraphics[width=2.5cm]{figs/promotion}
% http://bloggeate.com/wp-content/uploads/2011/04/como-promocionar-tu-blog.jpg
%\end{center}
%\label{fig:whatwedo}
%\end{figure}

%\begin{center}
%GSyC/LibreSoft's tasks: research, teaching, development, promotion of free software.

%\end{center}
%\end{frame}

\usebackgroundtemplate{}

%--------------------------------------------------------
\frame{
\maketitle
\begin{center}
\includegraphics[width=2cm]{format/libresoft-logo}
\hspace{0.5cm}
\includegraphics[width=5cm]{format/gsyc-urjc}
\vspace{0.5cm}
\includegraphics[width=3cm]{format/emadrid.png}
\end{center}
}

\end{document}